---
title: "Clover ETL Tutorial"
author: "Bill Zichos"
date: "August 19, 2015"
output:
  html_document:
    keep_md: yes
---

Before we get started...
```{r}
setwd("~/GitHub/Clover-ETL")
unzip("sample-data.zip")
```

## Orders

Process the *orders* data.  This first involves loading and formatting the data.  Since we only care about orders placed in 2014, we will filter prior to merging with other data sets.


1. Load orders.txt
```{r}
orders <- read.fwf(paste(getwd(), "/orders.txt", sep = ""), widths = c(10, 10, 8, 8, 10),header = FALSE, skip = 1, as.is = TRUE, col.names = c("orderId", "custId", "itemId", "qty", "transDate"), colClasses = c("character", "character", "character", "numeric", "character"), strip.white = TRUE)

# format the date so we can do some processing on it later
orders$transDate <- as.Date(orders$transDate, "%d.%m.%Y")
```

2. Filter orders for 2014
```{r include = FALSE}
# load the "lubridate" library for some special date-handling features
library("lubridate")
```
```{r}
orders.2014 <- orders[year(orders$transDate)==2014,]
```

## Customers

1. Load customers.txt
```{r}
customers <- read.csv(paste(getwd(), "/customers.txt", sep =""), sep = ";", header = TRUE, col.names = c("CUST_ID", "CUST_NAME", "STREET", "CITY", "STATE", "POST_CODE"))

customers$CUST_ID <- as.character(customers$CUST_ID)
```

2. Filter C* State Customers
```{r}
customers.c <- customers[customers$STATE %in% c("CA", "CO", "CT"),]
```

## Items

Excel files are a bit trickier.  In reality, I would ask the source to change to CSV or manually change myself, but I want to prove that Excel can be handled in R.  This requires the "gdata" R package library and the Perl libraries. 

1. Load items.xlsx
```{r include = FALSE}
library(gdata)
# installed the perl libraries from http://www.activestate.com/activeperl/downloads
installXLSXsupport(perl = "C:\\Perl64\\bin\\perl.exe")
```
```{r}
items <- read.xls(paste(getwd(), "/items.xlsx", sep = ""), sheet = 1, header = TRUE, perl = "C:\\Perl64\\bin\\perl.exe")

items$ID <- as.character(items$ID)
```

2. Exclude non-2-liter items.
```{r}
items.2Liter <- items[items$Unit=="2-Liter",]
```

## Combine the datasets
First merge Orders with Customers.  Then take the result and merge with Items.  The resulting dataset has 304 observations and 13 variables.
```{r}
df <- merge(orders.2014, customers.c, by.x = "custId", by.y = "CUST_ID")
df <- merge(df, items.2Liter, by.x = "itemId", by.y = "ID")

# calculate the sales price - Unit Price * Quantity
df$totalSales <- df$qty * df$UnitPrice
```

```{r include = FALSE}
library(dplyr)
```
```{r}
df <- select(df, custId, custName = CUST_NAME, totalQty = qty, totalSales)

group.by.customer <- group_by(df, custId, custName)

# summarize by customer
df.agg <- arrange(summarize(group.by.customer, totalQty = sum(totalQty), totalSales = sum(totalSales)), totalQty)

# limit to just those with 30 or more orders
best.customers <- df.agg[df.agg$totalQty>=30,]

# sort by order total sales
best.customers[order(best.customers$totalSales, decreasing = TRUE),]
```